Code for Breast Cancer Pain Study

### 2020-08-09 manuscript_code_20200809.Rmd
table updated

## Re-name variables
# colnames(cdmd.pvt)[which(grepl(pattern="surgery",colnames(cdmd.pvt))==T)]
colnames(dmd)[which(colnames(dmd)=="Usual.Occupation.f2")] = "Occupation"
colnames(dmd)[which(colnames(dmd)=="marital.status.f2")] = "Marital.status"
colnames(dmd)[which(colnames(dmd)=="Primary.Payer.f2")] = "Insurance.type"
colnames(dmd)[which(colnames(dmd)=="race.f")] = "Race"
colnames(dmd)[which(colnames(dmd)=="stage.f")] = "Cancer.stage"
colnames(dmd)[which(colnames(dmd)=="age.f")] = "Age"
colnames(dmd)[which(colnames(dmd)=="chemo.f")] = "Chemo"
colnames(dmd)[which(colnames(dmd)=="surgery.f")] = "Surgery"
colnames(dmd)[which(colnames(dmd)=="Tobacco.History")] = "Tobacco.history"
```

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Blight",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

# Summary of variables

## Table 1 (Base)

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    pval = compare_means(formula,data=final.data)
    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
  }
  return(tmp.sum)
}

@@

## Table 1 (Base + more biomarker variables)

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    pval = compare_means(formula,data=final.data)
    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
  }
  return(tmp.sum)
}

@@

## Scatter plots

@@

### Poverty

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

# par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

@@

pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Poverty,y=meandata$Pain,
     xlim=c(0,70),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Poverty", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

dev.off()

# par(mfrow=c(3,3))
pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
par(mar=c(4,4,2,2))
plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
# cbind(meandata$size,cex)

x = c(20,40,60,80)
y = c(1.5,2.5,4,6)
points(x=rep(0.4,length(x)),
       y=seq(0.8,0.4,length.out=length(x)),
       cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
text(x=rep(0.55,length(x)),
     y=seq(0.8,0.4,length.out=length(x)),
     labels=x,cex=1.5)
text(x=0.5,y=0.95,
     labels="Number of patients",cex=1.5)
dev.off()

@@

### Blight

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Blight, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Blight,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Blight", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

legend("bottomright",bty="n",cex=1.5,
       legend=c(paste("exp(coef) =",round(exp(p.glm$coefficients[2]),digits=2)),
                paste("(P-value < .001)"),
                paste("")))
dev.off()

@@

# Analysis

## Figure 2

# Significance

@@

## Table 2

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

## Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

@@

## Inflation model
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

@@

# Basic model + more biomarker variables

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

## Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table

@@

## Inflation model:
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

>>>>

## Re-name variables
# colnames(cdmd.pvt)[which(grepl(pattern="surgery",colnames(cdmd.pvt))==T)]
colnames(dmd)[which(colnames(dmd)=="Usual.Occupation.f2")] = "Occupation"
colnames(dmd)[which(colnames(dmd)=="marital.status.f2")] = "Marital.status"
colnames(dmd)[which(colnames(dmd)=="Primary.Payer.f2")] = "Insurance.type"
colnames(dmd)[which(colnames(dmd)=="race.f")] = "Race"
colnames(dmd)[which(colnames(dmd)=="stage.f")] = "Cancer.stage"
colnames(dmd)[which(colnames(dmd)=="age.f")] = "Age"
colnames(dmd)[which(colnames(dmd)=="chemo.f")] = "Chemo"
colnames(dmd)[which(colnames(dmd)=="surgery.f")] = "Surgery"
colnames(dmd)[which(colnames(dmd)=="Tobacco.History")] = "Tobacco.history"
colnames(dmd)[which(colnames(dmd)=="count.Pain")] = "Visits"
```

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Blight",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Visits",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

# Table1: Summary of variables

## Base

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x) %in% c("numeric","integer")) {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x) %in% c("numeric","integer")) {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x) %in% c("numeric","integer")) {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    pval = compare_means(formula,data=final.data)
    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
  }
  return(tmp.sum)
}

@@

## Base + more biomarker variables

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

# get_summary = function(j) {
#   x = final.data[,j]
#   cname = colnames(final.data)[j]
#   if (class(x)=="numeric") {
#     tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
#                     paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
#     rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
#   } else {
#     tmp.sum = t(t(summary(x)))
#     tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
#     rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
#   }
#   return(tmp.sum)
# }
# table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
# 
# get_mean = function(j) {
#   x = final.data[,j]
#   cname = colnames(final.data)[j]
#   formula = as.formula(paste("Pain ~",cname))
#   if (class(x)=="numeric") {
#     if (! cname == "Pain") {
#       fit = glm(formula, data = final.data, family = "poisson")
#       tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
#     } else {
#       tmp.sum = t(t(c(NA,NA)))
#     }
#   } else {
#     tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
#   paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
#   }
#   return(tmp.sum)
# }
# 
# get_pval = function(j) {
#   x = final.data[,j]
#   cname = colnames(final.data)[j]
#   formula = as.formula(paste("Pain ~",cname))
#   if (class(x)=="numeric") {
#     if (! cname == "Pain") {
#       fit = glm(formula, data = final.data, family = "poisson")
#       tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
#     } else {
#       tmp.sum = t(t(c(NA,NA)))
#     }
#   } else {
#     tmp.sum = t(t(rep(NA,length(levels(x)))))
#     pval = compare_means(formula,data=final.data)
#     tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
#   }
#   return(tmp.sum)
# }

@@

# Plots

@@

## Poverty

@@

```{r,echo=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

@@

# pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Poverty,y=meandata$Pain,
     xlim=c(0,70),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Poverty", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

# dev.off()

# par(mfrow=c(3,3))
# pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
# par(mar=c(4,4,2,2))
# plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
# # cbind(meandata$size,cex)
# 
# x = c(20,40,60,80)
# y = c(1.5,2.5,4,6)
# points(x=rep(0.4,length(x)),
#        y=seq(0.8,0.4,length.out=length(x)),
#        cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
# text(x=rep(0.55,length(x)),
#      y=seq(0.8,0.4,length.out=length(x)),
#      labels=x,cex=1.5)
# text(x=0.5,y=0.95,
#      labels="Number of patients",cex=1.5)
# dev.off()

@@

## Blight

@@

```{r,echo=FALSE}
p.glm = glm(Pain ~ Blight, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

# pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Blight,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Blight", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

legend("bottomright",bty="n",cex=1.5,
       legend=c(paste("exp(coef) =",round(exp(p.glm$coefficients[2]),digits=2)),
                paste("(P-value < .001)"),
                paste("")))
# dev.off()

@@

# Table 2: Regression model

## Base

@@

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age")
explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age")
explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

### Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)

vars.list = sapply(explvars,function(x) {
  x.level <- levels(final.data[,which(colnames(final.data)==x)])
  if (! is.null(x.level)) {
    x.level[1] <- paste(x.level[1],"(reference)")
    return(x.level)
  } else {
    return(x)
  }
})
tablevars = c("(Intercept)",unlist(vars.list)) # all variables that will be printed in the table

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("",gsub(pattern="[.]",replacement=" ",x=explvars)),
          rnames = tablevars,
          n.rgroup = c(1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

@@

### Inflation model
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("",gsub(pattern="[.]",replacement=" ",x=explvars)),
          rnames = tablevars,
          n.rgroup = c(1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

@@

## Base + more biomarker variables

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

### Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars,function(x) {
  x.level <- levels(final.data[,which(colnames(final.data)==x)])
  if (! is.null(x.level)) {
    x.level[1] <- paste(x.level[1],"(reference)")
    return(x.level)
  } else {
    return(x)
  }
})
tablevars = c("(Intercept)",unlist(vars.list)) # all variables that will be printed in the table

@@

### Inflation model:
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("",gsub(pattern="[.]",replacement=" ",x=explvars)),
          rnames = tablevars,
          n.rgroup = c(1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
